### 2018-03-07 latex/sc18/algorithm.tex
minor updates

Given a model $\T{M}=\CP$, we compute the relative error $\|\TA - \T{M}\|/\|\TA\|$ efficiently by using the identity $\|\TA-\T{M}\|^2 = \|\TA\|^2 - 2\langle \TA, \T{M} \rangle + \|\T{M}\|^2.$
The quantity $\|\TA\|$ is fixed, and the other two terms can be computed cheaply given the temporary matrices computed during the course of the BCD algorithm.
The second term can be computed using the identity $\langle \TA, \T{M} \rangle = \langle \Mn{M}{N}, \Mn{H}{N} \rangle$, where $\Mn{M}{N} = \Mz{A}{N} (\Mn{H}{N-1} \Khat \cdots \Khat \Mn{H}{1})$ is the MTTKRP result in the $N$th mode (the MTTKRP result is computed during the course of the algorithm to update the factor matrix).
The third term can be computed using the identity $\|\T{M}\|^2 = \V{1}^\Tra(\M{S} \Hada $

@@

\begin{algorithm}
\caption{$(\CPl,\epsilon) = \text{Par-NNCP}(\TA,k)$}
\label{alg:2D}
\begin{algorithmic}[1]
\Require $\TA$ is an $I_1\times \cdots \times I_N$ tensor distributed across a $P_1\times \cdots \times P_N$ grid of $P$ processors, so that $\TA_{\V{p}}$ is $(I_1/P_1)\times \cdots \times (I_N/P_N)$ and is owned by processor $\V{p}=(p_1,\dots,p_N)$, $k$ is rank of approximation
%\Require Local memory: space for $\Mn{H}{i}_{p_i}$ of size $(M_i/P_i) \times k$, for $1\leq i\leq N$
\State \Comment{Initialize data}
\State $a = \text{Norm-Squared}(\TA_{\V{p}})$
\State $\alpha = \text{All-Reduce}(a,\textsc{All-Procs})$
\For{$n=2$ to $N$}
	\State Initialize $\Mn{H}{n}_{\V{p}}$ of dimensions $(I_n/P)\times k$ 
	\State $\M{U} = \text{Local-SYRK}(\Mn{H}{n}_{\V{p}})$
	\State $\Mn{G}{n} = \text{All-Reduce}(\M{U},\textsc{All-Procs})$
	\State $\Mn{H}{n}_{p_n} = \text{All-Gather}(\Mn{H}{n}_{\V{p}},\textsc{Proc-Slice}(n,\VE{p}{n}))$
\EndFor
\State \Comment{Compute NNCP approximation}
\While{not converged}
	\State \Comment{Perform outer iteration of BCD}
	\For{$n=1$ to $N$}
	\State \Comment{Compute new factor matrix in $n$th mode}
	\State $\M{V} = \text{Local-MTTKRP}(\TA_{p_1\cdots p_N},\{\Mn{H}{i}_{p_i}\},n)$
		\label{line:locMTTKRP}
	\State $\Mn{M}{n}_{\V{p}} = \text{Reduce-Scatter}(\M{V},\textsc{Proc-Slice}(n,\VE{p}{n}))$ 
		\label{line:reduce-scatter}
	\State $\M{S} = \Mn{G}{1} \Hada \cdots \Hada \Mn{G}{n-1} \Hada \Mn{G}{n+1} \Hada \cdots \Hada \Mn{G}{N}$
		\label{line:hadamard}
	\State $\Mn[\hat]{H}{n}_{\V{p}} = \text{Local-NLS-Update}(\M{S},\Mn{M}{n}_{\V{p}})$
		\label{line:locNLS}
	\State \Comment{Normalize columns}
	\State $\V{c} = \text{Local-Col-Norms}(\Mn[\hat]{H}{n}_{\V{p}})$
	\State $\V{\lambda} = \text{All-Reduce}(\V{c},\textsc{All-Procs})$
	\State $\Mn{H}{n}_{\V{p}} = \text{Local-Col-Scale}(\Mn[\hat]{H}{n}_{\V{p}},\V{\lambda})$
	\State \Comment{Organize data for later modes}
	\State $\M{U} = \text{Local-SYRK}(\Mn{H}{n}_{\V{p}})$
		\label{line:locSYRK}
	\State $\Mn{G}{n} = \text{All-Reduce}(\M{U},\textsc{All-Procs})$
		\label{line:all-reduce}
	\State $\Mn{H}{n}_{p_n} = \text{All-Gather}(\Mn{H}{n}_{\V{p}},\textsc{Proc-Slice}(n,\VE{p}{n}))$
		\label{line:all-gather}
	\EndFor
	\State \Comment{Compute relative error $\epsilon$ from mode-$N$ matrices}
	\State $b = \text{Inner-Product}(\Mn{M}{N}_{\V{p}},\Mn[\hat]{H}{N}_{\V{p}})$
	\State $\beta = \text{All-Reduce}(b,\textsc{All-Procs})$
	\State $\gamma = \V{\lambda}^\Tra (\M{S} \Hada \Mn{G}{N}) \V{\lambda}$
	\State $\epsilon = \sqrt{(\alpha-2\beta+\gamma)/\alpha}$ 
\EndWhile
\Ensure $\|\TA - \dsquare{\V{\lambda}; \Mn{H}{1},\dots,\Mn{H}{N}}\| /\|\TA\| = \epsilon$
\Ensure Local matrices: $\Mn{H}{n}_{\V{p}}$ is $(I_n/P)\times k$ and owned by processor $\V{p}=(p_1,\dots,p_N)$, for $1\leq n \leq N$, $\V{\lambda}$ stored redundantly on every processor
\end{algorithmic}
\end{algorithm}

>>>>

Given a model $\T{M}=\CP$, we compute the relative error $\|\TA - \T{M}\|/\|\TA\|$ efficiently by using the identity $\|\TA-\T{M}\|^2 = \|\TA\|^2 - 2\langle \TA, \T{M} \rangle + \|\T{M}\|^2.$
The quantity $\|\TA\|$ is fixed, and the other two terms can be computed cheaply given the temporary matrices computed during the course of the BCD algorithm.
The second term can be computed using the identity $\langle \TA, \T{M} \rangle = \langle \Mn{M}{N}, \Mn{H}{N} \rangle$, where $\Mn{M}{N} = \Mz{A}{N} (\Mn{H}{N-1} \Khat \cdots \Khat \Mn{H}{1})$ is the MTTKRP result in the $N$th mode.
The third term can be computed using the identity $\|\T{M}\|^2 = \V{1}^\Tra(\Mn{S}{N} \Hada \MnTra{H}{N} \Mn{H}{N})\V{1}$ where $\Mn{S}{N}=\MnTra{H}{1} \Mn{H}{1} \Hada \cdots \Hada \MnTra{H}{N-1} \Mn{H}{N-1}$.
Both matrices $\Mn{M}{N}$ and $\Mn{S}{N}$ are computed during the course of the BCD algorithm for updating the factor matrix $\Mn{H}{N}$.

@@

\begin{algorithm}
\caption{$(\CPl,\epsilon) = \text{Par-NNCP}(\TA,k)$}
\label{alg:2D}
\begin{algorithmic}[1]
\Require $\TA$ is an $I_1\times \cdots \times I_N$ tensor distributed across a $P_1\times \cdots \times P_N$ grid of $P$ processors, so that $\TA_{\V{p}}$ is $(I_1/P_1)\times \cdots \times (I_N/P_N)$ and is owned by processor $\V{p}=(p_1,\dots,p_N)$, $k$ is rank of approximation
%\Require Local memory: space for $\Mn{H}{i}_{p_i}$ of size $(M_i/P_i) \times k$, for $1\leq i\leq N$
\State \Comment{Initialize data}
\State $a = \text{Norm-Squared}(\TA_{\V{p}})$
\State $\alpha = \text{All-Reduce}(a,\textsc{All-Procs})$
\For{$n=2$ to $N$}
	\State Initialize $\Mn{H}{n}_{\V{p}}$ of dimensions $(I_n/P)\times k$ 
	\State $\M{U} = \text{Local-SYRK}(\Mn{H}{n}_{\V{p}})$
	\State $\Mn{G}{n} = \text{All-Reduce}(\M{U},\textsc{All-Procs})$
	\State $\Mn{H}{n}_{p_n} = \text{All-Gather}(\Mn{H}{n}_{\V{p}},\textsc{Proc-Slice}(n,\VE{p}{n}))$
\EndFor
\State \Comment{Compute NNCP approximation}
\While{not converged}
	\State \Comment{Perform outer iteration of BCD}
	\For{$n=1$ to $N$}
	\State \Comment{Compute new factor matrix in $n$th mode}
	\State $\M{V} = \text{Local-MTTKRP}(\TA_{p_1\cdots p_N},\{\Mn{H}{i}_{p_i}\},n)$
		\label{line:locMTTKRP}
	\State $\Mn{M}{n}_{\V{p}} = \text{Reduce-Scatter}(\M{V},\textsc{Proc-Slice}(n,\VE{p}{n}))$ 
		\label{line:reduce-scatter}
	\State $\Mn{S}{N} = \Mn{G}{1} \Hada \cdots \Hada \Mn{G}{n-1} \Hada \Mn{G}{n+1} \Hada \cdots \Hada \Mn{G}{N}$
		\label{line:hadamard}
	\State $\Mn[\hat]{H}{n}_{\V{p}} = \text{Local-NLS-Update}(\Mn{S}{N},\Mn{M}{n}_{\V{p}})$
		\label{line:locNLS}
	\State \Comment{Normalize columns}
	\State $\V{c} = \text{Local-Col-Norms}(\Mn[\hat]{H}{n}_{\V{p}})$
	\State $\V{\lambda} = \text{All-Reduce}(\V{c},\textsc{All-Procs})$
	\State $\Mn{H}{n}_{\V{p}} = \text{Local-Col-Scale}(\Mn[\hat]{H}{n}_{\V{p}},\V{\lambda})$
	\State \Comment{Organize data for later modes}
	\State $\M{U} = \text{Local-SYRK}(\Mn{H}{n}_{\V{p}})$
		\label{line:locSYRK}
	\State $\Mn{G}{n} = \text{All-Reduce}(\M{U},\textsc{All-Procs})$
		\label{line:all-reduce}
	\State $\Mn{H}{n}_{p_n} = \text{All-Gather}(\Mn{H}{n}_{\V{p}},\textsc{Proc-Slice}(n,\VE{p}{n}))$
		\label{line:all-gather}
	\EndFor
	\State \Comment{Compute relative error $\epsilon$ from mode-$N$ matrices}
	\State $b = \text{Inner-Product}(\Mn{M}{N}_{\V{p}},\Mn[\hat]{H}{N}_{\V{p}})$
	\State $\beta = \text{All-Reduce}(b,\textsc{All-Procs})$
	\State $\gamma = \V{\lambda}^\Tra (\Mn{S}{N} \Hada \Mn{G}{N}) \V{\lambda}$
	\State $\epsilon = \sqrt{(\alpha-2\beta+\gamma)/\alpha}$ 
\EndWhile
\Ensure $\|\TA - \dsquare{\V{\lambda}; \Mn{H}{1},\dots,\Mn{H}{N}}\| /\|\TA\| = \epsilon$
\Ensure Local matrices: $\Mn{H}{n}_{\V{p}}$ is $(I_n/P)\times k$ and owned by processor $\V{p}=(p_1,\dots,p_N)$, for $1\leq n \leq N$, $\V{\lambda}$ stored redundantly on every processor
\end{algorithmic}
\end{algorithm}
